Exploring Local Explanations of Non-linear Models Using Animated Linear Projections

### 2022-02-03 _paper.rmd
Rotate fifa and ames use cases: basis below scatterplot; and knit.

<!-- What-if & counterfactual analysis -->
Using a radial tour can be considered similar to counterfactual, what-if analysis, such as _ceteris paribus_ [@biecek_ceterisparibus_2020]. This phrase, Latin for "other things held constant" or "all else unchanged", shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant. It ignores correlations of the features and imagines a case that was not observed. In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object. A constraint in our approach is that the basis must remain orthonormal. When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction. This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.

>>>>

<!-- What-if & counterfactual analysis -->
Using a radial tour can be compared with counterfactual, what-if analysis, such as _ceteris paribus_ profiles [@biecek_ceterisparibus_2020]. Ceteris paribus, is Latin for "other things held constant" or "all else unchanged". These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant. It ignores correlations of the features and imagines a case that was not observed. In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object. A constraint in our approach is that the basis must remain orthonormal. When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction. This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.

### 2022-02-03 _paper.tex
Rotate fifa and ames use cases: basis below scatterplot; and knit.

Using a radial tour can be considered similar to counterfactual, what-if analysis, such as \emph{ceteris paribus} (Biecek 2020). This phrase, Latin for ``other things held constant'' or ``all else unchanged,'' shows how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant. It ignores correlations of the features and imagines a case that was not observed. In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object. A constraint in our approach is that the basis must remain orthonormal. When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction. This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.

>>>>

Using a radial tour can be compared with counterfactual, what-if analysis, such as \emph{ceteris paribus} profiles (Biecek 2020). Ceteris paribus, is Latin for ``other things held constant'' or ``all else unchanged.'' These profiles show how an instance's prediction would change from a marginal change in one explanatory feature given that other features are held constant. It ignores correlations of the features and imagines a case that was not observed. In contrast, our approach is a geometric explanation of the factual; it varies contributions of the features by rotating the basis, a reorientation of the data object. A constraint in our approach is that the basis must remain orthonormal. When the contribution of one feature decreases, the contributions of others necessarily increase such that there is a complete component in that direction. This also ensures that what is seen is strictly a low-dimensional projection from high-dimensions and is thus an interpretable visualization.
